---
title: "Eating Out"
author: "Carmen Ramos-Caminos"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The data comes from the 1998 General Social Survey answer this question   

Is there a difference in frequency/times people eat out based on race?

$H_0$ There is no difference among the various races with frequency of eating out  
$H_A$ There is a statistical difference between at least two races with the frequency of eating out  

```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS1998.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,age,eatout)
mydata <- na.omit(mydata)
```

# Methods
Response variable is categorical:  dine out last week? yes or no

Explanatory variable is race. It is a categorical variable, so we use cat~cat. 

Inference: chi-squared and fisher exact tests 

# Descriptive summaries 

## Graphical summary 

### Barchart 

```{r}
barchartGC(~race + eatout,data=mydata)
barchartGC(~race + eatout,data=mydata, type="percent")

```
  
By looking at the frequency graph black people and others tend to eat out about the same in frequency. White individuals tend to eat out more often.  

## Numerical Summary  

```{r}
table1 <- xtabs(~race + eatout, data=mydata)
rowPerc(table1)
colPerc(table1)
```
  
In terms of eating out white people eat out 83% of the time, black people eat out 69% percent of time, and others eat out 72% of the time.  There is a significant difference between others and white individuals.  

# Results 

## Inferential results  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
  
Chi-squared test shows that the p-value is 0.000002287 much less than 0.05. Because the value is so small I reject the null hypothesis.  There is a difference in the frequency individuals eat out. In order to conclude where the difference is we must use the fisher exact test. 

```{r}
fisherXY(mydata$race,mydata$eatout, printAll = TRUE)
```
  
When comparing white individuals to black individuals the p-Value is too small due to difference in frequency of white and black individuals eating out. When comparing white and other there is a difference the p-value is so small that it the results did not happen by chance.  We do not have statistical evidence of difference between black and other groups. The odds ratio means is 2.2 to 1. That means that white people are twice as likely to eat out than black people. Whites are 1.9 more likely to eat out than others.


# Conclusion  

There is a statistical difference between at least two races in regard to the frequency of eating out.  I reject the null hypothesis the p-value is 0.000002287 much less than 0.05.  When comparing white to black individuals  whites tend to eat out more frequently with a ratio of  2.2:1 when compared to other individuals whites are more 1.9 times more likely to eat out. By using the Fisher exact test I am able to prove that whites are more likely to eat out more frequently than black and others. Therefore, with cetainly allowing me to  reject the null hypothesis.  








