---
title: "Eating Out"
author: "Carmen Ramos-Caminos"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The data comes from the 1998 General Social Survey answer this question   

Is there a difference in frequency/times people eat out based on race?

$H_0$ There is no difference among the various races with frequency of eating out  
$H_A$ There is a statistical difference between at least two races with the frequency of eating out  


```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS1998.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,age,eatout)
mydata <- na.omit(mydata)
```

# Methods
Response variable is categorical:  dine out last week? yes or no

Explanatory variable is race. It is a categorical variable, so we use cat~cat. 

Inference: chi-squared and fisher exact tests 

# Descriptive summaries 

## Graphical summary 

### Barchart 

```{r}
barchartGC(~race + eatout,data=mydata)
barchartGC(~race + eatout,data=mydata, type="percent")

```
  
By looking at the frequency graph black people and others tend to eat out about the same in frequency. White individuals tend to eat out more often.  

## Numerical Summary  

```{r}
table1 <- xtabs(~race + eatout, data=mydata)
rowPerc(table1)
colPerc(table1)
```
  
In terms of eating out white people ate out 83% of the time compared to black people at out 69% percent of time and others eat 72%.  

# Results 

## Inferential results  


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
  
Chi-squared test shows that the p-value is 0.000002287 much less than 0.05. Because the value is so small I reject the null hypothesis.  

```{r}
fisherXY(mydata$race,mydata$eatout, printAll = TRUE)
```
  
When comparing white people to black people the p-Value is too small. When comparing white and other there is a difference the p-value is so small that it the results did not happen by chance.  We do not have statistical evidence between black and other groups. The odds ratio means is 2.2 to 1. That means that they whites are twice as likely to eat out than blacks. Whites are 1.9 more likely to eat out than other.


# Conclusion  

 I reject the null hypothesis. There is a statistical difference between at least two races with the frequency of eating out.  








